---
title: "MTh - Interpretation of results: r/AutismInWomen corpus"
subtitle: "Continues from the topic modelling document (Step 16, k=6 final)"
author: "J. E."
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Context

This document continues directly from the end of the topic modelling
document (`step3_topicmodelling_reddit_lemmatizedcorpus.Rmd`).
Preprocessing, vocabulary pruning, and k selection (k=6, v2, domain
stopwords removed) are closed and are not revisited here. Step 17 of
that document already justified, on the
model's own behavior, why V1 is residual noise rather than a genuine
topic (probability below chance, no meaningful secondary weight,
5 dominant documents with no shared theme, spread across different flairs); that reasoning is
not repeated here, only assumed.

This document starts from three files saved at the end of the topic
modelling document, none of which contains post text:

- `modelo_final_btm.rds`: the final BTM model (k=6, v2), saved in Step 16.
- `topic_dominante_final.rds`: dominant topic per document, saved in
  Step 17.1 (no flair attached yet; that join happens here, as part of
  interpretation, not inherited from the previous document).
- `id_flair.csv`: post id and `post_flair`, saved in Step 18.

Only V2 through V6 (5 topics) are treated as substantively
interpretable; V1 is shown in the reference table for completeness but
excluded from the flair-based figures below, and never used to name a
theme. In the write-up (Step 4), V1-V6 are referred to as Topic 1
through Topic 6 (dropping the "V" prefix), so numbering in the running
text matches the tables/figures 1-for-1, with Topic 1 corresponding to
the residual, non-interpreted component.

**Note:** the flair of each post is read from `id_flair.csv` (post id and flair only), which avoids loading the full corpus, so this document runs without any post text. Table 1 and Figures 1 and 2 are saved in `outputs/`, together with the numbers behind Figure 2.

## Libraries

```{r libraries}
# Load packages
library(tidyverse)
library(tidytext)
library(BTM)
library(kableExtra)
```

```{r output-folders}
# Folders for the tables and figures saved by this document
dir.create("outputs/tables", recursive = TRUE, showWarnings = FALSE)
dir.create("outputs/figures", recursive = TRUE, showWarnings = FALSE)
```

## Loading the saved files

```{r load-files}
# Load the final model and the dominant topic per document
modelo_final_btm <- readRDS("modelo_final_btm.rds")
topic_dominante_final <- readRDS("topic_dominante_final.rds")

modelo_final_btm
nrow(topic_dominante_final)
```

## Step 1: Joining post_flair (external validation, done here)

```{r join-flair}
# Add each post's flair and count posts with no flair
validacion_flair_final <- topic_dominante_final |>
  rename(id = doc_id) |>
  left_join(read_csv("outputs/tables/id_flair.csv"), by = "id")

nrow(validacion_flair_final)
sum(is.na(validacion_flair_final$post_flair))
```

## Step 2: Top terms per topic

Reporting cutoff fixed at n=10, the usual convention for reporting
top terms.

```{r top-terms}
# Top 10 words per topic (function reused for the n = 15 check below)
resumen_topics <- function(modelo, top_n = 10) {
  terms_list <- terms(modelo, top_n = top_n)
  map_dfr(seq_along(terms_list), function(i) {
    tibble(topic = paste0("V", i), top_10_words = paste(terms_list[[i]]$token, collapse = ", "))
  })
}

top_terms <- resumen_topics(modelo_final_btm, top_n = 10)
top_terms
```

### Step 2b: Robustness check (n=15, not for reporting)

```{r top-terms-check}
# Top 15 words per topic, as a robustness check
top_terms_check <- resumen_topics(modelo_final_btm, top_n = 15) |>
  rename(top_15_words = top_10_words)

top_terms_check
```

## Step 3: Results section outputs (table + figures)

This section has three pieces: a table and two figures. No
`topic_labels` table is built here; the labels are given in prose in
Step 4.

### Table 1: Topic summary

```{r topic-summary-table}
# Table 1: top words and number of dominant documents per topic
n_dominant <- validacion_flair_final |>
  count(topic, name = "n_dominant_docs")

topic_summary <- top_terms |>
  left_join(n_dominant, by = "topic") |>
  mutate(status = if_else(topic == "V1", "Residual (not interpreted)", "Substantive"))

# Save the table (also shown below)
write_csv(topic_summary, "outputs/tables/table1_topic_summary.csv")

knitr::kable(topic_summary) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = TRUE,
    font_size = 13
  ) |>
  column_spec(2, width = "50%")
```

### Figure 1: Top terms per topic, V1 flagged as residual

```{r fig-top-terms}
# Figure 1: top 10 words per topic
top_terms_plot <- terms(modelo_final_btm, top_n = 10) |>
  imap_dfr(~ .x |> mutate(topic = paste0("V", .y))) |>
  mutate(
    is_residual = topic == "V1",
    token = reorder_within(token, probability, topic)
  )

fig_top_terms <- ggplot(top_terms_plot, aes(probability, token, fill = is_residual)) +
  geom_col(show.legend = TRUE) +
  facet_wrap(~ topic, scales = "free_y", ncol = 3) +
  scale_y_reordered() +
  scale_fill_manual(
    values = c(`TRUE` = "grey70", `FALSE` = "#2c7fb8"),
    labels = c(`TRUE` = "V1: residual, not interpreted", `FALSE` = "Substantive topic"),
    name = NULL
  ) +
  labs(title = "Top 10 terms per topic (BTM, k = 6)", x = "Probability", y = NULL) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

ggsave("outputs/figures/fig1_top_terms.png", fig_top_terms, width = 9, height = 6, dpi = 300)

fig_top_terms
```

### Figure 2: Topic composition within each flair (external validation)

This answers the more useful question for validation: given a flair a user
picked independently of the model, what share of posts with that flair
fall into each topic? V1 is excluded (5 dominant documents total,
<0.01% of the corpus, would be an invisible or noisy sliver in every
bar). Flairs are ordered by volume, largest first, since there are 10+
categories and an unordered axis would be hard to read.

This is evidence to keep in mind for Step 4 (labeling): any claim like
"topic X aligns with flair Y" needs to be checked against this actual
output.

#### Step 3b: Dropping flairs with too few posts before plotting

Flairs with very few posts (a handful of documents) produce
near-100%-of-one-color bars in the figure below that look like strong
signal but are actually sampling noise, not real concentration. These
are filtered out before plotting, not just visually ignored, so the
figure only shows flairs large enough to say something meaningful.

```{r flair-counts}
# Count posts per flair and keep flairs with at least 50 posts
flair_counts <- validacion_flair_final |>
  filter(!is.na(post_flair)) |>
  count(post_flair, name = "n_posts", sort = TRUE)

flair_counts

min_posts_flair <- 50

flairs_validos <- flair_counts |>
  filter(n_posts >= min_posts_flair) |>
  pull(post_flair)

n_distinct(flair_counts$post_flair)
length(flairs_validos)
```

5 flairs fall below the n=50 cutoff (13, 10, 7, 2, and 1 posts,
33 posts combined out of 57,041), all dropped. `LGBTQIA+` (n=75) stays
in, just above the line.

```{r fig-flair-stacked}
# Figure 2: topic composition within each flair
flair_order <- flair_counts |>
  filter(post_flair %in% flairs_validos) |>
  pull(post_flair)

flair_topic_prop <- validacion_flair_final |>
  filter(!is.na(post_flair), topic != "V1", post_flair %in% flairs_validos) |>
  mutate(post_flair = factor(post_flair, levels = flair_order))

fig_flair_stacked <- ggplot(flair_topic_prop, aes(x = post_flair, fill = topic)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Topic composition within each flair",
    subtitle = paste0("V1 (residual) and flairs with n < ", min_posts_flair, " posts excluded from this view"),
    x = NULL, y = "Proportion of posts", fill = "Topic"
  ) +
  theme_minimal(base_size = 9) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8))

ggsave("outputs/figures/fig2_flair_topic_composition.png", fig_flair_stacked, width = 9, height = 6, dpi = 300)

fig_flair_stacked
```

The numbers behind Figure 2 are saved as a table: for each flair, the percentage of its posts in each topic (V1 and the small flairs excluded, as in the figure).

```{r flair-topic-table}
# Percentage of each flair's posts in each topic (the numbers behind Figure 2)
flair_topic_table <- flair_topic_prop |>
  count(post_flair, topic) |>
  group_by(post_flair) |>
  mutate(pct = round(100 * n / sum(n), 1)) |>
  ungroup()

write_csv(flair_topic_table, "outputs/tables/fig2_flair_topic_percentages.csv")

flair_topic_table
```

## Step 4: Topic labeling

Evidence strength is uneven across topics and that unevenness is kept
visible below rather than smoothed into a uniform narrative: Topics 2
and 3 are backed by both word content and independent flair
concentration; Topics 4, 5, and 6 rest on word content alone, with no
flair that validates them externally.

**Topic 1: residual, not interpreted.** Not a labeling decision but a
non-decision, documented in the topic modelling document (Step 17): average topic
probability below chance (1/6), essentially no document giving it
meaningful secondary weight, and the 5 documents where it does
dominate share no theme and are spread across different flairs. Treated as noise
throughout, never named.

Topic 2 (*Relationships & social connection*): words *friend*,
*talk*, *relationship*, and *social* center on interpersonal
connection. This is confirmed by flair: the *Relationships* flair
(n = 2,200) has about 81% of its posts in Topic 2, the
strongest single piece of flair evidence in the model.

Topic 3 (*Diagnosis journey*): words *diagnose*, *diagnosis*, *adhd*,
and *experience* center on the diagnostic process and conditions that
co-occur with autism. The *Diagnosis Journey* flair (n = 1,076) is
similarly concentrated, about 83% of its posts fall into Topic 3.

Topic 4 (*Emotional struggle*): words *meltdown*, *anxiety*, *cry*,
and *bad* point to emotional distress, while words *struggle*,
*issue*, and *hard* frame that distress as an ongoing difficulty. No
flair in this corpus names this theme directly, so no external
validation was used.

Topic 5 (*Sensory sensibilities*): words *wear*, *eat*, *sensory*,
and *noise* read as a catalogue of sensory input across modalities,
with *sensory* itself among them. Words *hate* and *issue* suggest
this isn't purely descriptive, some of that input may be experienced
as difficult or unwanted. No external flair was used for this
topic's validation.

Topic 6 (*Daily life functionality*): words *job*, *school*, *home*,
and *week* center on the routine side of daily life. Work and school
refer to practical responsibilities, rather than the relational or
emotional side covered in other topics. No flair was used to validate
this externally either.

