---
title: "MTh - LDA topic modelling on the scientific literature corpus"
subtitle: "k selection, model fitting, and topic extraction"
author: "J. E."
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = TRUE)
```

## Introduction

This document builds the LDA topic model for the scientific literature corpus, starting from `tokens_clean_lemmatized_literature.csv` (16 papers, 58,060 tokens, 6,896 lemmas, unpruned).

k is selected by a qualitative comparison of the resulting topics.

**Note:** the table that summarises k selection (top terms and dominant papers for k = 4 to 8) is saved in `outputs/tables/`, together with the topic-term and paper-topic tables of the final model (Step 2.4). The seed is passed to `LDA()` through `control = list(seed = 123)`, so repeated runs give the same models.

```{r}
# Load packages
library(tidyverse)
library(tidytext)
library(topicmodels)
```

```{r}
# Folder for the tables saved by this document
dir.create("outputs/tables", recursive = TRUE, showWarnings = FALSE)
```

## Step 1: Building the document-term matrix

The tokens are not pruned by document frequency, so any noise still hiding in them survives straight into the DTM. Building the DTM is what surfaced it.

```{r}
# Read the token table and the full udpipe annotation from the preprocessing
tokens_clean <- read_csv("tokens_clean_lemmatized_literature.csv")
annotation   <- readRDS("annotation_literatura.rds")  # full udpipe annotation, used in 1.3

nrow(tokens_clean)
```

```{r}
# Number of papers
n_distinct(tokens_clean$paper_id)
```

```{r}
# Build a document-term matrix (papers x lemmas)
dtm <- tokens_clean |>
  count(paper_id, lemma) |>
  cast_dtm(document = paper_id, term = lemma, value = n)

dtm
```

`Maximal term length: 72` is the first red flag; no real English lemma is that long.

### 1.1: URL, email, and DOI fragments

```{r}
# The 5 longest lemmas, to spot noise
tokens_clean |>
  mutate(lemma_length = str_length(lemma)) |>
  arrange(desc(lemma_length)) |>
  distinct(lemma, lemma_length) |>
  head(5)
```

`udpipe` failed to split these into separate tokens, so they survived as one giant "word" each. A pattern-based filter removes them: URLs (`http`, `www.`), common domains (`.com`, `.org`, `.edu`, `.net`, `.gov`), email addresses (`@`), DOI prefixes, and any term containing `/` (which also catches article codes like `aut0010.1177/...`). This same pattern also catches ORCID identifiers, since they always appear as full `https://orcid.org/...` URLs in this corpus rather than as bare digit strings.

```{r}
# Remove URLs, email addresses and DOI fragments
tokens_clean_v2 <- tokens_clean |>
  filter(!str_detect(lemma, "http|www\\.|\\.(com|org|edu|net|gov)|@|^doi|/"))

nrow(tokens_clean)
```

```{r}
# Tokens after removing URL, email and DOI fragments
nrow(tokens_clean_v2)
```

### 1.2: Digits glued to words

```{r}
# Number of distinct lemmas that contain digits
tokens_clean_v2 |>
  filter(str_detect(lemma, "\\d")) |>
  distinct(lemma) |>
  nrow()
```

```{r}
# The 10 most frequent lemmas that contain digits
tokens_clean_v2 |>
  filter(str_detect(lemma, "\\d")) |>
  count(lemma, sort = TRUE) |>
  head(10)
```

314 terms: article codes, table labels, dates, citation numbers stuck to the preceding word (`pmc2025`, `table3`, `april2025`…). Before removing them blanket-style, worth checking that legitimate diagnostic-instrument names with numbers in them (DSM-5, AQ-10, RAADS-14…) aren't caught in the same net.

```{r}
# Check that instrument names such as DSM, ICD or SRS survive as separate tokens
tokens_clean_v2 |>
  filter(str_detect(lemma, regex("^(dsm|icd|aq|raads|srs|ados|adi)$", ignore_case = TRUE))) |>
  count(lemma, sort = TRUE)
```

They survive as separate tokens (`udpipe` split "DSM-5" into "dsm" and "5"), so it's safe to drop every remaining term with an embedded digit.

```{r}
# Remove lemmas with embedded digits
tokens_clean_v3 <- tokens_clean_v2 |>
  filter(!str_detect(lemma, "\\d"))

nrow(tokens_clean_v2)
```

```{r}
# Tokens after removing lemmas with digits
nrow(tokens_clean_v3)
```

### 1.3: Citation surnames tagged as proper nouns

The trickiest case. Author surnames cited in the running text (not just in the references block, which was already excluded in the EDA) get tagged `PROPN` and kept by the `upos` filter, for example "Chapman", "Sasson", "Guba".

```{r}
# Find lemmas tagged as proper nouns every time they appear
lemma_upos_summary <- annotation |>
  mutate(lemma_lower = str_to_lower(lemma)) |>
  filter(lemma_lower %in% tokens_clean_v3$lemma) |>
  distinct(lemma_lower, upos) |>
  group_by(lemma_lower) |>
  summarise(tagged_propn_only = all(upos == "PROPN"), .groups = "drop")

propn_only_terms <- lemma_upos_summary |> filter(tagged_propn_only)
nrow(propn_only_terms)
```

```{r}
# Share of tokens (%) in those lemmas
total_tokens_affected <- tokens_clean_v3 |> filter(lemma %in% propn_only_terms$lemma_lower) |> nrow()
round(100 * total_tokens_affected / nrow(tokens_clean_v3), 2)
```

904 terms flagged this way (words tagged `PROPN` every single time they appear, never as a common word elsewhere), 3.77% of the corpus by volume. Two automated rules were tried to separate real surnames from topically-relevant words that happened to get mistagged:

- **Surname + nearby year** (e.g. "Chapman… 2022"): worked for some citation styles but missed others, since numbered-citation papers never put the surname next to a year in the body text.
- **Always-PROPN + low document-frequency**: didn't hold up. Important terms like `symptoms` and `neuropsychiatric` turned out to have `doc_frequency == 1`, identical to genuine surnames like `sasson` or `heidegger`. There's no clean signal separating them.

```{r}
# In how many papers does each of those lemmas appear?
propn_doc_freq <- tokens_clean_v3 |>
  filter(lemma %in% propn_only_terms$lemma_lower) |>
  distinct(paper_id, lemma) |>
  count(lemma, name = "doc_frequency")

table(propn_doc_freq$doc_frequency)
```

797 of the 904 terms (88%) appear in only one document. For those, there's no reliable way to tell noise from rare-but-real content, so they're left untouched. The remaining 107 (`doc_frequency >= 2`) are a small enough list to review by hand instead.

```{r}
# Lemmas that appear in 2 or more papers (reviewed by hand)
propn_review_list <- propn_doc_freq |> filter(doc_frequency >= 2) |> arrange(desc(doc_frequency))
nrow(propn_review_list)
```

Manual review of those 107 confirmed 96 as real noise (surnames, first names, institutions, journals/publishers, countries, cities, months) and kept the rest as legitimate vocabulary (`traits`, `skills`, `faculty`, `lab`, `nursing`, `revised`, `spss`, `stata`, and `adi`, the ADI-R diagnostic instrument, same category as `dsm`/`srs`/`icd` confirmed safe earlier).

```{r}
propn_remove_list <- c(
  # surnames / first names
  "baron", "jones", "nicolaidis", "ritvo", "wilson", "allison", "begeer",
  "hayward", "kentrou", "mcdonnell", "milton", "napolitano", "pellicano",
  "wood", "ahmed", "baldwin", "belcher", "bertilsdotter", "brede", "costley",
  "delucia", "downie", "dubreucq", "fletcher", "fortuna", "garcía",
  "gillberg", "gray", "gruber", "hedlund", "howard", "hus", "jackson",
  "kanfiszer", "khudiakova", "kim", "kopp", "mclean", "mello", "miller",
  "milner", "moseley", "perry", "pisula", "raz", "rosqvist", "schneid",
  "strang", "szatmari", "taylor", "trundle", "vries", "williams", "sarah",
  "alexandra", "gemma", "hannah", "mary", "miranda", "danielle", "george",
  # institutions / journals / publishers / countries / cities / dates
  "commons", "united", "institute", "international", "australia", "england",
  "ireland", "london", "usa", "canada", "city", "birmingham", "basel",
  "italy", "netherlands", "north", "northern", "poland", "sweden",
  "switzerland", "los", "francis", "autistica", "pubmed", "scopus", "mdpi",
  "licensee", "inc.", "microsoft", "u.s.", "january", "february",
  "november", "december", "excellence"
)

tokens_clean_v4 <- tokens_clean_v3 |>
  filter(!lemma %in% propn_remove_list)

nrow(tokens_clean_v3)
```

```{r}
# Tokens after removing the reviewed proper nouns
nrow(tokens_clean_v4)
```

### 1.4: Final DTM

```{r}
# Final document-term matrix
dtm <- tokens_clean_v4 |>
  count(paper_id, lemma) |>
  cast_dtm(document = paper_id, term = lemma, value = n)

dtm
```

```{r}
# Save the document-term matrix (word counts per paper: kept local)
saveRDS(dtm, "dtm_literatura_final.rds")
```

16 documents by 6,438 terms, 84% sparsity. This is the input for k selection.

## Step 2: Selecting k

### 2.1: Log-likelihood doesn't help

```{r}
# Fit LDA for k = 2 to 10 and compare log-likelihood
set.seed(123)
k_candidates <- 2:10

lda_models <- map(k_candidates, ~ LDA(dtm, k = .x, control = list(seed = 123)))
names(lda_models) <- k_candidates

map_dbl(lda_models, logLik)
```

Log-likelihood keeps improving as k grows; it always will, since more topics can always fit the training data a bit better. It cannot pick k, so k is chosen by reading the topics.

### 2.2: Why k=2 is ruled out

k=2 is the smallest value in the range, so it is checked first: its top terms show whether two distinct themes emerge.

```{r}
# Function that lists the top 10 terms of each topic; top terms for k = 2
print_top_terms <- function(lda_model, n_terms = 10) {
  tidy(lda_model, matrix = "beta") |>
    group_by(topic) |>
    slice_max(beta, n = n_terms) |>
    arrange(topic, desc(beta)) |>
    summarise(top_terms = paste(term, collapse = ", "), .groups = "drop")
}

print_top_terms(lda_models[["2"]])
```

The two topics share 7 of their top 10 words ("autistic", "woman", "experience", "autism", "study", "social", "participant"). k=2 isn't finding two themes, it's splitting the corpus's shared vocabulary roughly in half. Ruled out.

### 2.3: Comparing k=4 through k=8

Two things checked for each candidate: whether the topics read as genuinely distinct (not two versions of the same theme), and whether they're reasonably spread across documents rather than dominated by one or two papers. That's a real risk with only 16 documents, where a model with, say, 8 topics is averaging just 2 papers per topic.

```{r}
# For k = 4 to 8: top terms and number of dominant papers per topic
for (k in 4:8) {
  cat("\n==================== k =", k, "====================\n")

  topics_k <- print_top_terms(lda_models[[as.character(k)]])
  for (i in 1:nrow(topics_k)) cat("Topic", topics_k$topic[i], ":", topics_k$top_terms[i], "\n")

  gamma_k <- tidy(lda_models[[as.character(k)]], matrix = "gamma")
  doc_spread <- gamma_k |> group_by(topic) |> summarise(n_docs_dominant = sum(gamma > 0.5), .groups = "drop")
  cat("\nDocuments per topic (gamma > 0.5):", paste(doc_spread$n_docs_dominant, collapse = ", "), "\n")
}
```

```{r}
# Top terms and number of dominant papers per topic for k = 4 to 8, saved as a table
k_comparison <- map_dfr(4:8, function(k) {
  model_k <- lda_models[[as.character(k)]]
  spread <- tidy(model_k, matrix = "gamma") |>
    group_by(topic) |>
    summarise(n_docs_dominant = sum(gamma > 0.5), .groups = "drop")
  print_top_terms(model_k) |>
    left_join(spread, by = "topic") |>
    mutate(k = k, .before = topic)
})

write_csv(k_comparison, "outputs/tables/lda_k4_k8_topics.csv")
```

Reading through the range:

- **k=4 and k=5** both keep suicidal ideation and flourishing/wellbeing mixed into a single topic, two fairly opposite themes stuck together.
- **k=6** is the first point where those two separate into their own topics, and every topic is still shared by at least 2 documents.
- **k=7 and k=8** separate things a bit further, but a cited author's surname ("hampton") shows up among one topic's top words, and some topics become dominated by a single document. Those are signs the model is starting to fit individual papers rather than themes shared across the corpus.

### 2.4: Decision

k=6 is the only point in the range compared (k = 4 to 8) where topics read as genuinely distinct *and* no topic is dominated by a single paper. k is chosen by this qualitative read, plus one extra check, document spread, that matters especially with a corpus this small.

The fitted model is saved locally. Its two probability tables, topic-term (`beta`) and paper-topic (`gamma`), are saved as CSV; they are what the interpretation of the topics needs.

```{r}
# Final model (k = 6): save the model object locally, and its two tables as CSV
lda_final <- lda_models[["6"]]
saveRDS(lda_final, "lda_final_k6.rds")

tidy(lda_final, matrix = "beta")  |> write_csv("outputs/tables/lda_final_beta.csv")
tidy(lda_final, matrix = "gamma") |> write_csv("outputs/tables/lda_final_gamma.csv")
```
