---
title: "MTh - Interpretation of results: scientific literature corpus"
subtitle: "Continues from the LDA topic modelling document (k=6 final)"
author: "J. E."
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Context

This document continues directly from the end of the LDA topic
modelling document. The EDA, preprocessing, and k selection are closed
and are not revisited here; this document starts from the final LDA
model (k=6) and focuses on interpretation: naming topics from their top
words and identifying representative papers per topic.

Methodology: evidence before assumption, each interpretive claim backed
by top words, document-level evidence, or the papers' own titles, not
by intuition alone.

`autistic`/`autism` are treated as anchor terms throughout: present
near the top of every topic by corpus design (the inclusion criterion
for all 16 papers), not because they carry discriminating signal. They
are kept in the model and in every table/figure below (removing them
would mean rebuilding the DTM and re-running k selection, out of scope
here); they are simply not used to name topics, and are visually
flagged as anchors in Figure 1.

**Note:** the interpretation runs from the two tables saved by the LDA document, without the model object or any article text. Tables and figures are saved in `outputs/`.

## Libraries

```{r libraries}
# Load packages
library(tidyverse)
library(tidytext)
library(ggplot2)
library(kableExtra)
```

```{r output-folders}
# Folders for the tables and figures saved by this document
dir.create("outputs/tables", recursive = TRUE, showWarnings = FALSE)
dir.create("outputs/figures", recursive = TRUE, showWarnings = FALSE)
```

## Loading the final model tables

This document starts from two tables saved at the end of the LDA topic
modelling document (Step 2.4): `lda_final_beta.csv` (probability of each
term in each topic) and `lda_final_gamma.csv` (share of each topic in
each paper).

```{r load-tables}
# Load the topic-term (beta) and paper-topic (gamma) tables of the final model
beta_tbl  <- read_csv("outputs/tables/lda_final_beta.csv")
gamma_tbl <- read_csv("outputs/tables/lda_final_gamma.csv")

count(beta_tbl, topic)
n_distinct(gamma_tbl$document)
```

## Step 1: Top terms per topic

Reporting cutoff fixed at n=10, as in Step 2.3 of the LDA document and
the usual convention for LDA reporting.

```{r top-terms}
# Top 10 terms per topic
top_terms <- beta_tbl |>
  group_by(topic) |>
  slice_max(beta, n = 10) |>
  arrange(topic, desc(beta)) |>
  summarise(top_terms = paste(term, collapse = ", "), .groups = "drop")

# Save the table (also shown below)
write_csv(top_terms, "outputs/tables/lda_top_terms_n10.csv")

knitr::kable(top_terms) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 12)
```

### Step 1b: Robustness check (n=15, not for reporting)

Extending past the n=10 cutoff only to check whether noise (citation
surnames, structural artifacts like "manuscript", "table", "figure",
"doi", "supplementary") shows up just past the reporting window. If it
does, that topic gets labeled more cautiously even though its top-10
looks clean. This table is a validation step, not something that goes
in the TFM.

Topic 3 shows `hampton` (a cited author's surname) and
`manuscript` at positions 14-15, past the reporting cutoff. Label that
topic with extra caution.

```{r top-terms-check}
# Top 15 terms per topic, as a robustness check
top_terms_check <- beta_tbl |>
  group_by(topic) |>
  slice_max(beta, n = 15) |>
  arrange(topic, desc(beta)) |>
  summarise(top_terms_15 = paste(term, collapse = ", "), .groups = "drop")

knitr::kable(top_terms_check) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 12)
```

### Figure 1: Top terms per topic, anchors highlighted

```{r fig-top-terms}
# Figure 1: top 10 terms per topic
top_terms_plot <- beta_tbl |>
  group_by(topic) |>
  slice_max(beta, n = 10) |>
  ungroup() |>
  mutate(
    is_anchor = term %in% c("autistic", "autism"),
    term = reorder_within(term, beta, topic)
  )

fig_top_terms <- ggplot(top_terms_plot, aes(beta, term, fill = is_anchor)) +
  geom_col(show.legend = TRUE) +
  facet_wrap(~ topic, scales = "free_y", ncol = 3) +
  scale_y_reordered() +
  scale_fill_manual(
    values = c(`TRUE` = "grey70", `FALSE` = "#2c7fb8"),
    labels = c(`TRUE` = "Anchor term (present across all topics)", `FALSE` = "Discriminating term"),
    name = NULL
  ) +
  labs(title = "Top 10 terms per topic (LDA, k = 6)", x = expression(beta), y = NULL) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

ggsave("outputs/figures/fig1_top_terms.png", fig_top_terms, width = 9, height = 6, dpi = 300)

fig_top_terms
```

## Step 2: Document-topic assignment (gamma)

Which papers dominate each topic (`gamma > 0.5`), to anchor labels in
actual documents rather than in word lists alone.

Note: gamma values in this corpus are extreme (almost all > 0.95, most
> 0.9999), meaning topic membership is close to binary rather than
mixed, likely a consequence of the small corpus (16 papers, 6 topics).
The one exception is paper 4 (gamma = 0.787 in topic 1), which shows
some genuine mixture; worth a closer look before treating it as a
clean topic-1 case.

```{r gamma-assignment}
# Dominant topic of each paper
doc_topic <- gamma_tbl |>
  group_by(document) |>
  slice_max(gamma, n = 1, with_ties = FALSE) |>
  arrange(topic, desc(gamma))

# Save the table (also printed below)
write_csv(doc_topic, "outputs/tables/lda_doc_topic.csv")

doc_topic
```

### Figure 2: Topic membership per paper (gamma, stacked)

```{r fig-gamma-stacked}
# Figure 2: topic membership per paper
gamma_matrix <- gamma_tbl |>
  mutate(
    document = factor(as.integer(document), levels = sort(unique(as.integer(document)))),
    topic = factor(topic)
  )

fig_gamma_stacked <- ggplot(gamma_matrix, aes(x = document, y = gamma, fill = topic)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Topic membership per paper (gamma), LDA k = 6",

    x = "Paper ID", y = "Proportion of topic membership", fill = "Topic"
  ) +
  theme_minimal(base_size = 11)

ggsave("outputs/figures/fig2_gamma_stacked.png", fig_gamma_stacked, width = 9, height = 6, dpi = 300)

fig_gamma_stacked
```

## Reference: Paper titles (supporting evidence)

Full titles for the 16 papers in the final corpus, used below as
supporting evidence alongside gamma and top words. The analysis text of
each paper is its title, abstract and body, so the words of the title
were part of the input to the model, together with the rest of the
paper. A match between a title and a topic's top words is therefore a
consistency check, not independent validation: it shows that the topic
reflects what the paper says about itself. A title is a small part of
the text (6 to 28 words in a text of roughly 4,000 to 13,000), so it is
used here to identify what each dominant paper is about.

```{r paper-titles}
# Full titles of the 16 papers (supporting evidence)
paper_titles <- tribble(
  ~paper_id, ~title,
  1,  "Overlapping And Differentiating Clinical Features of Autism and Borderline Personality Disorder in Women and People Assigned Female at Birth: A Cross-Sectional Study",
  2,  "Navigating Residual Diagnostic Categories: The Lived Experiences of Women Diagnosed with Autism and ADHD in Adulthood",
  3,  "Disability Barriers Autistic Girls Face in Secondary Education: A Systematic Review",
  4,  "The Healthcare Experiences of Middle and Older Age Autistic Women in the United Kingdom",
  5,  "Autistic Women's Experiences of Interpersonal Violence",
  6,  "A Good Autistic Life: An Autistic-Led Conceptualization of Autistic Flourishing Through Autistic Women's Lived Experiences",
  7,  "Autistic Women's Maternal Health: A Review of the Literature on Prenatal, Birth, and Postpartum Experiences for Autistic Mothers",
  8,  "Autistic Women Report High Levels of Pain, Pain-Related Anxiety, and Pain Interference",
  9,  "The Impact of Intersectional Disadvantage on Autistic Women's Experiences of Interpersonal Violence: A Narrative Review",
  11, "\"I ask them what autism means for them\": A Qualitative Study of Staff Experiences of Working with Autistic Women and Birthing People in Community Perinatal Mental Health Teams",
  12, "Comparative Analysis of Autistic Women Across the Lifespan: Childhood vs. Adulthood Diagnosis",
  13, "Underdiagnosed and Misunderstood: Clinical Challenges and Educational Needs of Healthcare Professionals in Identifying Autism Spectrum Disorder in Women",
  14, "The Association Between Autism, Camouflaging and Anxiety with Suicidal Ideation in Women",
  15, "Diagnostic Challenges of Autism Spectrum Disorder in Women Without Intellectual or Language Impairments: A Narrative Review",
  16, "Eating Behaviors of Autistic Women with an Eating Disorder",
  17, "'I Fit the Category of the Box, It Just Doesn't Describe Me Well.' Exploring the Perspectives of Autistic Women and Gender-Diverse Individuals on Self-Report Autism Measures"
)

knitr::kable(paper_titles) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 12)
```

## Step 3: Topic labeling

Evidence strength varies across topics, kept visible rather than
smoothed over. There is no external label for the papers in this
corpus, so paper titles are used as supporting evidence: a title is a
short summary of the paper's subject, and when it matches a topic's top
words it supports the label. Because the title is part of the text the
model used, this is a consistency check, not independent validation.

`autistic`/`autism` are anchor terms throughout, present by corpus
design, not used to justify any label below.

**Topic 1: Autistic women across the lifespan: healthcare &
diagnosis.** Top words (experience, social, study, participant, girl,
woman, bpd, healthcare, school, research) span three different papers
rather than one clean theme. Two of the three titles point to a shared
thread: paper 3 is about autistic girls in secondary school, paper 4
about older autistic women in healthcare, together suggesting age as a
connecting idea across the lifespan. The third, paper 1, is about
telling autism apart from BPD, unrelated to age; it's grouped here by
shared vocabulary, not a shared topic. The label names the age pattern
where it actually holds (2 of 3 papers), not all three. Paper 4 is
also the only paper in the whole corpus with real topic mixture rather
than a clean assignment, worth keeping in mind.

**Topic 2: Diagnostic processes & health support systems.** Top words
(diagnose, participant, adhd, health, diagnosis, support, research,
mental, individual, service) are supported by titles across the topic,
though from two angles rather than one. Paper 2 (ADHD/autism diagnosis
in adulthood) and paper 12 (childhood vs. adulthood diagnosis
comparison) are squarely about the diagnostic process itself. Paper
11, titled around healthcare staff's experience in perinatal mental
health teams, sits on the other side of the same system: diagnosis
doesn't happen in isolation, it's mediated by healthcare professionals
and support services, which is plausibly where
`mental`/`service`/`support` come from.

**Topic 3: Mental health risk and female autism experiences.** Top
words (suicidal, ideation, anxiety, camouflaging, non-autistic,
mother) match paper 14's title directly (camouflaging, anxiety, and
suicidal ideation). Paper 7, the other dominant paper, is titled
around prenatal, birth, and postpartum experiences in autistic women,
which is where `mother` comes from. The connection between the two
papers may be precisely that: the experiences paper 7 addresses and
the symptoms or experiences paper 14 refers to.

**Topic 4: Diagnostic criteria challenges and gender differences.**
Top words (diagnostic, social, gender, female, participant,
experience, trait, study, individual, measure) are supported by
three dominant titles. Paper 15 and paper 13 are squarely about
diagnostic challenges for women. Paper 17, on perspectives around
self-report autism measures, fits the same theme from another angle:
self-report instruments are themselves part of the diagnostic process,
and a paper exploring their limits for autistic women and gender-
diverse people. The gender angle across all three plausibly points to
the same underlying issue: diagnostic criteria and measurement tools
that may or may not account well enough for how autism presents
differently across genders.

**Topic 5: Trauma experiences.** Top words (violence, wbp, behavior,
eat, interpersonal, report); `wbp` was confirmed as "women and birthing
people" by checking the source text directly. All three dominant
papers' titles support the topic directly: paper 5 and paper 9 are
both explicitly about autistic women's experiences of interpersonal
violence, and paper 16 is explicitly about eating behaviors/disorders
in autistic women. Interpersonal violence and disordered eating are
treated here as two expressions of the same underlying category,
trauma.

**Topic 6: Wellbeing and adversity symptoms.** Top words (pain,
anxiety, flourishing, describe, feel, neurotypical, study, woman,
life, analysis, social, related) can be read around how autistic
women experience life day to day, in relation to the world around
them. `social` and `neurotypical` may suggest a relational framing:
experience isn't described in isolation, but in contrast with others.
`pain`, `anxiety`, `feel`, and `life` may point to the everyday reality
of that experience; at the same time, `flourishing` can appear as one
possibility within that reality.

## Reference: Topic labels (final summary)

Manual table, not derived from any calculation, this is the outcome of
the labeling discussion above (top words + paper titles),
provided here as a quick-reference summary after the full reasoning
per topic.

```{r topic-labels-table}
# Manual table with the label of each topic
topic_labels <- tribble(
  ~topic, ~label,
  1, "Autistic women across the lifespan: healthcare & diagnosis",
  2, "Diagnostic processes & health support systems",
  3, "Mental health risk and female autism experiences",
  4, "Diagnostic criteria challenges and gender differences",
  5, "Trauma experiences",
  6, "Wellbeing and adversity symptoms"
)

# Save the table (also shown below)
write_csv(topic_labels, "outputs/tables/lda_topic_labels.csv")

knitr::kable(topic_labels, col.names = c("Topic", "Label")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 13)
```
